A Particle Swarm Optimisation-Based Indirect Approach to Web Service Composition

### 2015-10-26 paper.tex
Modified algorithms to be in plain English.

\begin{algorithm}[h]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
	\textbf{1.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
			\textbf{2.} Calculate the particle's fitness.\\
			\eIf{fitness value better than pBest}{
				\textbf{3a.} Assign current fitness as new pBest.\\
			}{
				\textbf{3b.} Keep previous pBest.\\
			}		
		}
		\textbf{4.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{5.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{6.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
	}

@@

\begin{algorithm}[h]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
 	\textcolor{blue}{\textbf{1.} Map each candidate service to an index in the particle's position vector.}\\
	\textbf{2.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
		    \textcolor{blue}{\textbf{3.} Create queue of services using the particle's position vector.}\\
			\textcolor{blue}{\textbf{4.} Build the corresponding composition graph using the queue.}\\
			\textcolor{blue}{\textbf{5.} Calculate the fitness of the resulting graph.}\\
			\eIf{fitness value better than pBest}{
				\textbf{6a.} Assign current fitness as new pBest.\\
			}{
				\textbf{6b.} Keep previous pBest.\\
			}		
		}
		\textbf{7.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{8.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{9.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
	}

@@

\begin{algorithm}[h]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{connectNode}{connectNode}\SetKwFunction{findCands}{findCands}\SetKwFunction{removeDangling}{removeDangling}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$I$, $O$, $queue$}
 \Output{composition graph $G$}
 $start.outputs \leftarrow \{I\}$\;
 $end.inputs \leftarrow \{O\}$\;
 $G.edges \leftarrow \{\}$\;
 $G.nodes \leftarrow \{start\}$\;
 $availOutputs \leftarrow \{start.outputs\}$\;
 $index \leftarrow 0$\;
 \While{$end.inputs \not\sqsubseteq availInputs$}{\label{buildingLine}
 $cand \leftarrow queue.get(index)$\;
 $index \leftarrow index + 1$\;
 \If{$cand.inputs \sqsubseteq availOutputs$}{
 \connectNode{$cand, G$}\;
 $currEndInputs \leftarrow availOutputs \cup \{cand.outputs\}$\;
 $queue \leftarrow queue.remove(index)$\;
 $index \leftarrow 0$\;
 }}
 \connectNode{$end, G$}\;
 \removeDangling{$G$}\;
 \KwRet $G$\;
 \vspace{2mm}
 \SetKwProg{myproc}{Procedure}{}{}
 \myproc{\connectNode{$n, G$}}{
   $inputsToFulfil \leftarrow \{cand.inputs\}$\;
 \While{$|inputsToFulfil| > 0$}{
	$graphN \leftarrow G.nodes.next()$\; 
	\If{$|n.inputs \sqcap graphN.outputs| > 0$}{
	  $inputsToFulfil \leftarrow inputsToFulfil - (n.inputs \sqcap graphN.outputs)$\;
	  $G.edges \leftarrow G.edges \cup \{graphN \rightarrow n\}$\;
    }
    $G.nodes \leftarrow G.nodes \cup \{n\}$\;
 }}
 \caption{Generating a composition graph from a queue.}
\label{graph_building}
\end{algorithm}

 As shown in Algorithm \ref{graph_building}, the values initially required are the composition task inputs ($I$), task outputs ($O$), and a $queue$ of services as its input; this leads to the creation of a composition graph $G$. We begin by assigning $I$ as the set of outputs produced by the graph's $start$ node, and $O$ as the set of input nodes required by the graph's $end$ node. Then, $start$ is added to the graph $G$, its outputs are added to a set ($availOutputs$) that records all available outputs from the nodes currently in the graph, and an $index$ variable is created to track positions in the $queue$ (holding 0, which indicates the queue head). After this setup stage, the following steps are repeated until $availOutputs$ can be used to fulfil all of the inputs of $end$: the node at the current $queue$ position ($index$) is retrieved as a candidate ($cand$); if all of its inputs can be fulfilled, $cand$ is connected to the graph using $connectNode$ -- which works by identifying existing nodes in the graph whose output fulfils the input of $cand$, and creating connecting edges from these existing nodes to $cand$ --, its outputs are added to $availOutputs$, it is removed from the queue, and $index$ is reset to 0; otherwise, the candidate in the next queue position is considered. Once all $end$ node inputs can be fulfilled, $end$ is connected to the graph, any dangling nodes are removed, and $G$ is returned. In summary, by using this algorithm the structure of the resulting graph changes depending on the service ordering within the queue provided as input.

@@

\begin{algorithm}[h]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
 	\textcolor{blue}{\textbf{1.} Identify layers and map each candidate service to an index in the particle's position vector.}\\
	\textbf{2.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
		    \textcolor{blue}{\textbf{3.} Decode and calculate quality of corresponding solution layer by layer.}\\
			\eIf{fitness value better than pBest}{
				\textbf{4a.} Assign current fitness as new pBest.\\
			}{
				\textbf{4b.} Keep previous pBest.\\
			}		
		}
		\textbf{5.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{6.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{7.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
		\Indm \textcolor{blue}{\textbf{8.} Create final solution graph based using gBest's weights.}\\
	}

@@

The unique feature of the PSO approach proposed in this section is that it identifies the \textit{composition layer} to which a service belongs. Before the optimisation process begins, the service repository is run through a discovery process \cite{wang2013genetic} that identifies the services that could be possibly used in the composition. As shown in Algorithm \ref{discovery}, this filtering process requires the set of inputs ($I$) and the set of outputs ($O$) from the overall composition task, in addition to the service repository ($R$); given these inputs, it produces a list of candidate service layers that are relevant to the composition ($L$). The algorithm keeps track of all available inputs so far ($C_{search}$), and uses them to discover additional layers: if a previously undiscovered service has all of its inputs satisfied by $C_{search}$, then it is added to the current layer and its outputs are added as available inputs in $C_{search}$. The discovery continues until no additional layers are found, and the final step verifies whether the desired composite output $O$ can in fact be achieved using the services in the repository. Once the composition layers are identified, the particle mapping takes place. As shown in in Figure \ref{layeredMapping}, each service layer is mapped to contiguous particle indices, effectively segmenting particles according to the layers. This segmentation facilitates the solution decoding process discussed in the following subsection.

\begin{algorithm}
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \LinesNumbered
 \SetNlSty{}{}{:}
 %\SetLine % For v3.9
 \Input{$I$, $O$, $R$}
 \Output{service layers $L$}
 $C_{search} \leftarrow I$\;
 $L \leftarrow \{\}$\;
 $i \leftarrow i$\;
 $S_{found} \leftarrow DiscoverService()$\;
 \While{$|S_{found}| > 0$}{
  $L.add(S_{found},i)$\;
  $i \leftarrow i + 1$\;
  $C_{search} \leftarrow C_{search} \cup C_{output}$ of $S_{found}$\;
  $S_{found} \leftarrow DiscoverService()$\;
  }
  \eIf{$C_{search} \supseteq O$}{
   \Return $L$\;
   }{
   Report no solution\;
 }
\caption{Discovering relevant service composition layers \cite{wang2013genetic}.}
\label{discovery}
\end{algorithm}

@@

\begin{algorithm}
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{filterByLayer}{filterByLayer}\SetKwFunction{mergeLayers}{mergeLayers}\SetKwFunction{findHighestTime}{findHighestTime}
 \SetKwFunction{eNull}{null}\SetKwFunction{getInputsSatisfied}{getInputsSatisfied}\SetKwFunction{calculateFitness}{calculateFitness}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$end$, sorted layers $L$}
 \Output{fitness $f$}
 $solution \leftarrow \{\}$\;
 $cost \leftarrow 0$\;
 $availability \leftarrow 1$\;
 $reliability \leftarrow 1$\;
 $nextToSatisfy \leftarrow \{\}$\;
 \ForAll{$inputs$ of $end$}{
    $nextToSatisfy \leftarrow nextToSatisfy \cup \{(input, end.time, |L|)\}$\;
 }
 $i \leftarrow |L|$\;
 \For{$i=|L|$ to $i=1$}{
    $toSatisfy \leftarrow \filterByLayer(nextToSatisfy,i)$\;
    $nextToSatisfy \leftarrow nextToSatisfy - toSatisfy$\;
    $mergedL \leftarrow \mergeLayers(L,i)$\;
    \While{$|toSatisfy| > 0$}{
	$nextS \leftarrow mergedL.next()$\;
	\eIf{$nextS = \eNull$}{
	    $nextToSatisfy \leftarrow nextToSatisfy \cup toSatisfy$\;
	    $toSatisfy \leftarrow \{\}$\;
	}{
	    $nLayer \leftarrow nextS.layer$\;
	    $satisfied \leftarrow \getInputsSatisfied(toSatisfy,nextS)$\;
	    \If{$|satisfied| > 0$}{
		$solution \leftarrow solution \cup \{nextS\}$\;
		$cost \leftarrow cost + nextS.QoS.time$\;
		$availability \leftarrow availability \times nextS.QoS.availability$\;
		$reliability \leftarrow reliability \times nextS.QoS.reliability$\;
	    }
	    $time \leftarrow nextS.QoS.time$\;
	    $toSatisfy \leftarrow toSatisfy - satisfied$\;
	    $highestTime \leftarrow \findHighestTime(satisfied)$\;
	    \ForAll{$inputs$ of $nextS$}{
		$nextToSatisfy \leftarrow nextToSatisfy \cup \{(input, time + highestTime, nLayer)\}$\;
	    }
	}
    }
 }
 $time \leftarrow \findHighestTime(nextToSatisfy)$\;
 $f \leftarrow \calculateFitness(cost,time,availability,reliability)$\;
 \Return $f$\;
\caption{Algorithm for decoding solutions and calculating their fitness.}
\label{solutionDecoding}
\end{algorithm}

@@

\begin{algorithm}
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{filterByLayer}{filterByLayer}\SetKwFunction{mergeLayers}{mergeLayers}
 \SetKwFunction{eNull}{null}\SetKwFunction{getInputsSatisfied}{getInputsSatisfied}\SetKwFunction{createEdges}{createEdges}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$start$, $end$, sorted layers $L$}
 \Output{final graph $G$}
 $G.nodes \leftarrow \{end\}$\;
 $G.edges \leftarrow \{\}$\;
 $nextToSatisfy \leftarrow \{\}$\;
 \ForAll{$inputs$ of $end$}{
    $nextToSatisfy \leftarrow nextToSatisfy \cup \{(input, end, |L|)\}$\;
 }
 $i \leftarrow |L|$\;
 \For{$i=|L|$ to $i=1$}{
    $toSatisfy \leftarrow \filterByLayer(nextToSatisfy,i)$\;
    $nextToSatisfy \leftarrow nextToSatisfy - toSatisfy$\;
    $mergedL \leftarrow \mergeLayers(L,i)$\;
    \While{$|toSatisfy| > 0$}{
	$nextS \leftarrow mergedL.next()$\;
	\eIf{$nextS = \eNull$}{
	    $nextToSatisfy \leftarrow nextToSatisfy \cup toSatisfy$\;
	    $toSatisfy \leftarrow \{\}$\;
	}{
	    $nLayer \leftarrow nextS.layer$\;
	    $satisfied \leftarrow \getInputsSatisfied(toSatisfy,nextS)$\;
	    \If{$|satisfied| > 0$}{
		$G.nodes \leftarrow G.nodes \cup \{nextS\}$\;
		$G.edges \leftarrow G.edges \cup \createEdges(nextS, satisfied)$\;
		$toSatisfy \leftarrow toSatisfy - satisfied$\;
		\ForAll{$inputs$ of $nextS$}{
		    $nextToSatisfy \leftarrow nextToSatisfy \cup \{(input, nextS, nLayer)\}$\;
		}
	    }
	}
    }
 }
 $G.nodes \leftarrow G.nodes \cup \{start\}$\;
 $G.edges \leftarrow G.edges \cup \createEdges(start,nextToSatisfy)$\;
 \Return $G$\;
\caption{Algorithm for building final graph solution.}
\label{solutionBuilding}
\end{algorithm}

>>>>

\begin{algorithm}[!htb]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
	\textbf{1.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
			\textbf{2.} Calculate the particle's fitness.\\
			\eIf{fitness value better than pBest}{
				\textbf{3a.} Assign current fitness as new pBest.\\
			}{
				\textbf{3b.} Keep previous pBest.\\
			}		
		}
		\textbf{4.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{5.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{6.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
	}

@@

\begin{algorithm}[!htb]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
 	\textcolor{blue}{\textbf{1.} Map each candidate service to an index in the particle's position vector.}\\
	\textbf{2.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
		    \textcolor{blue}{\textbf{3.} Create queue of services using the particle's position vector.}\\
			\textcolor{blue}{\textbf{4.} Build the corresponding composition graph using the queue.}\\
			\textcolor{blue}{\textbf{5.} Calculate the fitness of the resulting graph.}\\
			\eIf{fitness value better than pBest}{
				\textbf{6a.} Assign current fitness as new pBest.\\
			}{
				\textbf{6b.} Keep previous pBest.\\
			}		
		}
		\textbf{7.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{8.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{9.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
	}

@@

\begin{algorithm}[!htb]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{connectNode}{connectNode}\SetKwFunction{findCands}{findCands}\SetKwFunction{removeDangling}{removeDangling}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$I$, $O$, $queue$}
 \Output{composition graph $G$}
 Create $start$ node with outputs $I$ and $end$ node with inputs $O$\;
 Create graph $G$ containing the $start$ node\;
 Create set of available outputs containing $start$ outputs\;
 \While{available outputs do not satisfy $end$ inputs}{
 Get next candidate from queue\;
 \If{candidate inputs are satisfied by available outputs}{
 Connect node to graph\;
 Remove it from queue and go back to the queue's beginning\;
 }}
 Connect end node\;
 Remove dangling nodes from graph $G$\;
 \KwRet $G$\;
 \caption{Generating a composition graph from a queue}
\label{graph_building}
\end{algorithm}

 As shown in Algorithm \ref{graph_building}, the values initially required are the composition task inputs ($I$), task outputs ($O$), and a $queue$ of services as its input; this leads to the creation of a composition graph $G$. We begin by assigning $I$ as the set of outputs produced by the graph's $start$ node, and $O$ as the set of input nodes required by the graph's $end$ node. Then, $start$ is added to the graph $G$, its outputs are added to a set that records all available outputs from the nodes currently in the graph. After this setup stage, the following steps are repeated until the available outputs can be used to fulfil all of the inputs of $end$: the next node of the queue is retrieved as a candidate; if all of its inputs can be fulfilled, the candidate is connected to the graph, its outputs are added to the set of available outputs, and it is removed from the queue; otherwise, the candidate in the next queue position is considered. Once all $end$ node inputs can be fulfilled, $end$ is connected to the graph, any dangling nodes are removed, and $G$ is returned. In summary, by using this algorithm the structure of the resulting graph changes depending on the service ordering within the queue provided as input.

@@

\begin{algorithm}[!htb]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
 	\textcolor{blue}{\textbf{1.} Identify layers and map each candidate service to an index in the particle's position vector.}\\
	\textbf{2.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
		    \textcolor{blue}{\textbf{3.} Decode and calculate quality of corresponding solution layer by layer.}\\
			\eIf{fitness value better than pBest}{
				\textbf{4a.} Assign current fitness as new pBest.\\
			}{
				\textbf{4b.} Keep previous pBest.\\
			}		
		}
		\textbf{5.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{6.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{7.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
		\Indm \textcolor{blue}{\textbf{8.} Create final solution graph based using gBest's weights.}\\
	}

@@

The unique feature of the PSO approach proposed in this section is that it identifies the \textit{composition layer} to which a service belongs. Before the optimisation process begins, the service repository is run through a discovery process \cite{wang2013genetic} that identifies the services that could be possibly used in the composition. As shown in Algorithm \ref{discovery}, this filtering process requires the set of inputs ($I$) and the set of outputs ($O$) from the overall composition task, in addition to the service repository ($R$); given these inputs, it produces a list of candidate service layers that are relevant to the composition ($L$). The algorithm keeps track of all available outputs so far, and uses them to discover additional layers: if a previously undiscovered service has all of its inputs satisfied by the set of available outputs, then it is added to the current layer and its outputs are added to the set of available outputs. The discovery continues until no additional layers are found, and the final step verifies whether the desired composite output $O$ can in fact be achieved using the services in the repository. Once the composition layers are identified, the particle mapping takes place. As shown in in Figure \ref{layeredMapping}, each service layer is mapped to contiguous particle indices, effectively segmenting particles according to the layers. This segmentation facilitates the solution decoding process discussed in the following subsection.

\begin{algorithm}[!htb]
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \LinesNumbered
 \SetNlSty{}{}{:}
 %\SetLine % For v3.9
 \Input{$I$, $O$, $R$}
 \Output{service layers $L$}
 Initialise output set with $I$\;
 Discover services satisfied by output set\;
 \While{at least one service discovered}{
 Add services as the next layer in $L$\;
 Add the outputs of these services to the output set\;
 Discover additional services satisfied by the updated output set\;
 }
  \eIf{Output set satisfies $O$}{
   \Return $L$\;
   }{
   Report no solution\;
 }
\caption{Discovering relevant service composition layers \cite{wang2013genetic}.}
\label{discovery}
\end{algorithm}

@@

\begin{algorithm}[!htb]
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{filterByLayer}{filterByLayer}\SetKwFunction{mergeLayers}{mergeLayers}\SetKwFunction{findHighestTime}{findHighestTime}
 \SetKwFunction{eNull}{null}\SetKwFunction{getInputsSatisfied}{getInputsSatisfied}\SetKwFunction{calculateFitness}{calculateFitness}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$end$, sorted layers $L$}
 \Output{fitness $f$}
 Initialise variables to keep track of QoS: $cost=0, availability=1, reliability=1$\;
 Set all inputs of $end$ as the next to satisfy, associating each input with time $0$ and layer $|L|+1$\;
 \For{all sorted layers in L, from $end$ to first}{
 	Identify the inputs from next-to-satisfy set that correspond to services in this layer\;
    \While{not all of these inputs have been satisfied}{
	Find the next service from the previous layers with the highest weight\;
	\If{service outputs satisfy at least one input}{
		Update running QoS values with service QoS (add to running cost, multiply with running availability and reliability)\;
		Add the inputs of service to the set of next inputs to satisfy, each associated with (service time + highest time from satisfied inputs) and current layer position\;
    }
 }
 Find the total composition time as the highest from the remaining set of next inputs to satisfy\;
 Calculate fitness $f$ using total QoS values\;
 \Return $f$\;
\caption{Algorithm for decoding solutions and calculating their fitness.}
\label{solutionDecoding}
\end{algorithm}

@@

\begin{algorithm}[!htb]
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{filterByLayer}{filterByLayer}\SetKwFunction{mergeLayers}{mergeLayers}
 \SetKwFunction{eNull}{null}\SetKwFunction{getInputsSatisfied}{getInputsSatisfied}\SetKwFunction{createEdges}{createEdges}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$start$, $end$, sorted layers $L$}
 \Output{final graph $G$}
 Create graph $G$ containing the end node\;
 Set all inputs of $end$ as the next to satisfy, associating each input with the $end$ node and layer $|L|+1$\;
 \ForAll{sorted layers in $L$, from $end$ to first}{
 	Identify the inputs from next-to-satisfy set that correspond to services in this layer\;
    \While{not all of these inputs have been satisfied}{
    	Find the next service from the previous layer with the highest weight\;
	    \If{the outputs of this service satisfy at least one of the inputs for this layer}{
			Add service node to graph\;
			Add edges connecting this service node to the nodes whose inputs it satisfies\;
			Add the inputs of the service node to the set of next inputs to satisfy, each associated with the service node and current layer position\;
 }}}
 Add $start$ node to graph\;
 Add edges connecting $start$ node the associated nodes of all remaining inputs in the next-to-satisfy set\;
 \Return $G$\;
\caption{Algorithm for building final graph solution.}
\label{solutionBuilding}
\end{algorithm}
